Providing a Unified Software Environment for Canada’s National Advanced Computing Centers

### 2019-04-05 paper.tex
Corrections

%
% The "author" command and its associated commands are used to define the authors and their affiliations.
% Of note is the shared affiliation of the first two authors, and the "authornote" and "authornotemark" commands
% used to denote shared contribution to the research.
\author{Maxime Boissonneault}
\email{maxime.boissonneault@calculquebec.ca}
\affiliation{%
  \institution{Université Laval, Calcul Québec, Compute Canada}
  \streetaddress{2325 Rue de l'Université}
  \city{Québec}
  \state{Québec}
  \country{Canada}
  \postcode{G1V 0A6}
}
\author{Bart Oldeman}
\email{bart.oldeman@calculquebec.ca}
\affiliation{%
  \institution{McGill University, Calcul Québec, Compute Canada}
  \streetaddress{845 Rue Sherbrooke Ouest}
  \city{Montréal}
  \state{Québec}
  \country{Canada}
  \postcode{H3A 0G4}
}

@@

\section{Distribution mechanism}
\label{sec:Distribution_mechanism}
A shared file system such as NFS, Lustre, or GPFS, can provide access to software packages on a single cluster. However, experience in the field of high performance computing shows that parallel file systems are often challenging to operate at scale in a reliable and performant manner, especially for the high I/O operations per second caused by application startups. In addition, a shared file system is a single point of failure for a cluster, which nearly all jobs rely on in order to run. An alternative for a single cluster is to install binary packages on each node, using a mechanism such as RPM. However, this can result in divergent configurations across nodes, may be limited by available space on the local disk, and requires an additional configuration management layer to keep the packages synchronized.

@@

\section{A layered approach}
\label{sec:Layered_environment}
To install software packages for our common environment, we follow a layered approach, as illustrated in Table~\ref{tab:layers}. First, we require a basic Linux installation. Any version of Linux with a kernel 2.6.32 or more recent should work. This layer is managed by the system administrators, and should contain any driver, such as GPU drivers or drivers for the network fabric, kernel modules, as well as privileged binaries.

@@

The third layer is the scientific application layer. It contains mostly scientific or performance critical applications, including for example MPI implementations, CUDA, mathematical libraries, and domain specific applications.

\begin{figure*}
  \includegraphics[width=0.7\textwidth]{layers.png}
  \caption{Layered structure of our software stack. The bottom layer (red) is entirely managed by the system administrators of each cluster. The second layer (grey) contains libraries that can be installed centrally, through Nix, but can be overridden locally. The third layer (yellow) contains everything that one would typically find in the Linux user space. The fourth layer (green) contains packages that you would typically find in an operating system, but for which multiple versions are required. They are installed through Nix, but wrapped into a module by EasyBuild. The top layer (blue) contains packages that are not typically installed in a Linux distribution, and are installed via EasyBuild.}
  \label{fig:layers}
\end{figure*}

@@

EasyBuild originated from the University of Ghent, but now has over 200 contributors from around the world. The framework is based on scripted recipes (in Python) to install packages, usually from source code. It is designed to create optimized binaries, and eases recompilation of multiple versions of a given application using many combinations of compilers or libraries. EasyBuild is composed of three parts: the framework, the EasyBlocks, and the EasyConfigs.

@@

EasyBlocks are Python classes that define the steps required to install a given software package. There are generick EasyBlocks, such as {\it ConfigureMake} or {\it CMakeMake}, as well as custom EasyBlocks for packages that have more complicated installation procedures, such as OpenFOAM or PETSc.

EasyConfigs files parametrize the execution of the code from the EasyBlocks. They define the software name and version, the URL from which the source files can be downloaded as well as options to be used for the various steps.

In most cases, there is already an EasyBlock for a given software package, or it can be installed easily with one of the generic EasyBlocks. In these cases, writing a new recipe is as simple as editing a configuration file. EasyBuild also supports command line options to reuse and existing EasyConfig file, while changing the software version or the toolchain being used.

Toolchains are a core concept in EasyBuild. They represent a set or subset of a compiler, an MPI implementation, mathematical libraries and CUDA. Having the ability to generate multiple builds of a given software simply by providing a command line argument to change the toolchain has proven unvaluable to our workflow, and has allowed us to recompile hundreds of packages in matter of days when deploying a new version of compiler or a cluster with a new CPU architecture.

>>>>

%
% The "author" command and its associated commands are used to define the authors and their affiliations.
% Of note is the shared affiliation of the first two authors, and the "authornote" and "authornotemark" commands
% used to denote shared contribution to the research.
\author{Maxime Boissonneault}
\email{maxime.boissonneault@calculquebec.ca}
\affiliation{%
  \institution{Université Laval, Calcul Québec, Compute Canada}
  \streetaddress{2325 Rue de l'Université}
  \city{Québec}
  \state{Québec}
  \country{Canada}
  \postcode{G1V 0A6}
}
\author{Bart E. Oldeman}
\email{bart.oldeman@calculquebec.ca}
\affiliation{%
  \institution{McGill University, Calcul Québec, Compute Canada}
  \streetaddress{845 Rue Sherbrooke Ouest}
  \city{Montréal}
  \state{Québec}
  \country{Canada}
  \postcode{H3A 0G4}
}

@@

\section{Distribution mechanism}
\label{sec:Distribution_mechanism}
A shared file system such as NFS, Lustre, or GPFS can provide access to software packages on a single cluster. However, experience in the field of high performance computing shows that parallel file systems are often challenging to operate at scale in a reliable and performant manner, especially for the high I/O operations per second caused by application startups. In addition, a shared file system is a single point of failure for a cluster, which nearly all jobs rely on in order to run. An alternative for a single cluster is to install binary packages on each node, using a mechanism such as RPM. However, this can result in divergent configurations across nodes, may be limited by available space on the local disk, and requires an additional configuration management layer to keep the packages synchronized.

@@

\section{A layered approach}
\label{sec:Layered_environment}
To install software packages for our common environment, we follow a layered approach, as illustrated in Table~\ref{tab:layers}. First, we require a basic Linux installation. Any version of Linux with a kernel 2.6.32 (as used by RHEL 6 and its derivatives) or more recent should work. This layer is managed by the system administrators, and contains all drivers, typically provided by kernel modules and closely associated libraries, such as GPU and network drivers, as well as privileged binaries.

@@

The third layer is the scientific application layer. It contains mostly scientific or performance critical applications, including for example MPI implementations, the CUDA toolkit, mathematical libraries, and domain specific applications.

\begin{figure*}
  \includegraphics[width=0.7\textwidth]{layers.png}
  \caption{Layered structure of our software stack. The bottom layer (red) is entirely managed by the system administrators of each cluster. The second layer (grey) contains libraries that can be installed centrally, through Nix, but can be overridden locally. The third layer (yellow) contains everything that one would typically find in the Linux user space. The fourth layer (green) contains packages that you would typically find in an operating system, but for which multiple versions are required. They are installed through Nix, but wrapped into a module by EasyBuild. The top layer (blue) contains packages that are not typically installed in a Linux distribution, and are installed via EasyBuild.}
  \label{tab:layers}
\end{figure*}

@@

EasyBuild originated from the University of Ghent, but now has over 200 contributors from around the world. The framework is based on scripted recipes (in Python) to install packages, usually from source code. It is designed to create optimized binaries, and eases recompilation of multiple versions of a given application using many combinations of compilers or libraries. EasyBuild is composed of three parts: the framework, easyblocks, and easyconfig files (also known as easyconfigs).

@@

Easyblocks are Python classes that define the steps required to install a given software package. There are generic easyblocks, such as {\it ConfigureMake} or {\it CMakeMake}, as well as custom easyblocks for packages that have more complicated installation procedures, such as OpenFOAM or PETSc.

Easyconfig files parametrize the execution of the code from the easyblocks. They define the software name and version, the URL from which the source files can be downloaded as well as options to be used for the various steps.

In most cases, there is already an easyblock for a given software package, or it can be installed easily with one of the generic easyblocks. In these cases, writing a new recipe is as simple as editing a configuration file. EasyBuild also supports command line options to reuse an existing easyconfig file, while changing the software version or the toolchain being used.

Toolchains are a core concept in EasyBuild. They represent a set or subset of a compiler, an MPI implementation, mathematical libraries and CUDA. Having the ability to generate multiple builds of a given software simply by providing a command line argument to change the toolchain has proven invaluable to our workflow, and has allowed us to recompile hundreds of packages in matter of days when deploying a new version of compiler or a cluster with a new CPU architecture.
